### 2015-07-15 thesis/thesis/abstract.tex
changes

In the last few years, communities of services have been studied in a certain numbers of proposals as virtual pockets of similar expertise. The motivation is to provide these services with high chance of discovery through better visibility, and to enhance their capabilities when it comes to provide requested functionalities. There are a number of proposed mechanisms and models on aggregating web services and making them cooperate within their communities. However, forming optimal and stable communities as coalitions to maximize individual and group efficiency and income for all the involved parties has not been addressed yet. Also, in the proposed frameworks of these communities, a common assumption is that residing services, which are supposed to be autonomous and intelligent, are competing over received requests. However, those services can also exhibit cooperative behaviors, for instance in terms of substituting each other. When competitive and cooperative behaviors and strategies are combined, autonomous services are said to be ``coopetitive''. Deciding to compete or cooperate inside communities is a problem yet to be investigated.

In this thesis, we first identify the problem of defining efficient algorithms for coalition formation mechanisms within communities and propose some results using cooperative game-theoretic techniques. We propose a mechanism for community membership requests and selections of web services in the scenarios where there is interaction between one community and many web services and scenarios where web services can join multiple established communities. Then in order to address the coopetitive relation within our web services, we propose a decision making mechanism for our web services to efficiently choose competition or cooperation strategies to maximize their payoffs. We prove that the proposed decision mechanism is
efficient and can be implemented in time linear in the length of the time period considered for the analysis and the number of services in the community. Moreover, we conduct extensive
simulations, analyze various scenarios, and confirm the obtained theoretical results using parameters from a real web services dataset.

>>>>

In the last few years, communities of services have been studied in a certain numbers of proposals as virtual pockets of similar expertise. The motivation is to provide these services with high chance of discovery through better visibility, and to enhance their capabilities when it comes to provide requested functionalities. There are a number of proposed mechanisms and models on aggregating web services and making them cooperate within their communities. However, forming optimal and stable communities as coalitions to maximize individual and group efficiency and income for all the involved parties has not been addressed yet. Moreover, in the proposed frameworks of these communities, a common assumption is that residing services, which are supposed to be autonomous and intelligent, are competing over received requests. However, those services can also exhibit cooperative behaviors, for instance in terms of substituting each other. When competitive and cooperative behaviors and strategies are combined, autonomous services are said to be ``coopetitive''. Deciding to compete or cooperate inside communities is a problem yet to be investigated.

In this thesis, we first identify the problem of defining efficient algorithms for coalition formation mechanisms. We study the community formation problem in two different settings: 1) Communities with centralized manager having complete information using cooperative game-theoretic techniques 2) Communities with distributed decision making mechanisms having incomplete information using training methods. We propose mechanisms for community membership requests and selections of web services in the scenarios where there is interaction between one community and many web services and scenarios where web services can join multiple established communities. Then in order to address the coopetitive relation within communities of web services, we propose a decision making mechanism for our web services to efficiently choose competition or cooperation strategies to maximize their payoffs. We prove that the proposed decision mechanism is efficient and can be implemented in time linear in the length of the time period considered for the analysis and the number of services in the community. Moreover, we conduct extensive simulations, analyze various scenarios, and confirm the obtained theoretical results using parameters from a real web services dataset.

### 2015-07-15 thesis/thesis/chap1/chap1.tex
changes

Web service communities are dynamic by design
\cite{DBLP:journals/ijebr/MaamarSTBB09}. In these communities, web
services are modeled as intelligent autonomous agents, where they
can adopt a strategy maximizing their payoff at any time. A web
service can ask joining a community and has the right to leave it.
Community managers can invite or ask a web service to leave in
order to maximize the community profit. Users can simply stop
sending requests to a web service which is not providing
satisfactory services. Thus, it is important to consider all the
parties involved in the decision making process about the
community formation and management. Most of the recent work on communities of
services are either user-centric and focus on user satisfaction
\cite{Chun02user-centricperformance} or system-centric and focus
on the whole system throughput, performance and utilization. There
are many contributions in distributed, grid, cluster and cloud
services which are system-centric. However, in real world
environments and applications, both users and service providers
are self-interested agents, aiming to maximize their own profit.
In those environments, both parties (users and services) will
collaborate as long as they are getting more benefits and payoff.
Our initial research question is: \emph{How can we model the
community of agent-based services in order to maximize the utility
of involved users, web services and community organizers? [R1]}

@@

\subsection{Contribution 2: Distributed decision making for dynamic formation of web services communities.}

@@

%Web service communities are dynamic by design

>>>>

Web services communities are dynamic by design
\cite{DBLP:journals/ijebr/MaamarSTBB09}. In these communities, web
services are modeled as intelligent autonomous agents, where they
can adopt a strategy maximizing their payoff at any time. A web
service can ask joining a community and has the right to leave it.
Community managers can invite or ask a web service to leave in
order to maximize the community profit. Users can simply stop
sending requests to a web service which is not providing
satisfactory services. Thus, it is important to consider all the
parties involved in the decision making process about the
community formation and management. Most of the recent work on communities of
services are either user-centric and focus on user satisfaction
\cite{Chun02user-centricperformance} or system-centric and focus
on the whole system throughput, performance and utilization. There
are many contributions in distributed, grid, cluster and cloud
services which are system-centric. However, in real world
environments and applications, both users and service providers
are self-interested agents, aiming to maximize their own profit.
In those environments, both parties (users and services) will
collaborate as long as they are getting more benefits and payoff.
Our initial research question is: \emph{How can we model the
community of agent-based services in order to maximize the utility
of involved users, web services and community organizers? [R1]}

@@

\subsection{Contribution 2: Distributed Decision Making for Dynamic Formation of Web Services Communities}

@@

%Web services communities are dynamic by design

### 2015-07-15 thesis/thesis/chap2/chap2.tex
changes

\subsection{Web Service Communities}\label{sec:CWSDefinition}
Community refers to ``the condition of sharing or having certain
attitudes and interests in common'' or ``a group of people living
in the same place or having a particular characteristic in
common''\footnote{Oxford Dictionaries}. In
\cite{DBLP:journals/internet/BenatallahSD03,
Zeng:2003:QDW:775152.775211}, the authors introduce community of
web services as collection of cooperative web services with common
functionalitiers but different QoS metrics. Therefore
$communities$ are differentiated from $composition$ types of web
service cooperation in which web services with different
functionalities work together to generate a new service with
composite functionality.

@@

Here we introduce the related research work regarding the engineering and formation
of communities of web services. In \cite{DBLP:journals/internet/BenatallahSD03}, Benatallah et al.
defined communities as \emph{Service Containers} that aggregate
substitutable web services providing a common functionality (same
set of operations). They abstracted \emph{Service Containers} as
web services that are created, advertised, discovered and invoked
just as elementary web services. The \emph{Container} is
considered as a manager that is responsible for web service
selection upon receiving a request on run-time. The authors have
proposed a scoring service based on non-functional requirements of
the request and web service capabilities to dynamically chose the
web service to perform the requested task. A similar concept was
proposed by Maamar et al. in
\cite{DBLP:journals/ijebr/MaamarSTBB09}. The authors introduced
web service communities as a collection of web services with a
common functionality but different QoS properties. A community
manager, upon receiving a request, delegates the request to one of
its current members. The choice is based on the performance
history and quality metrics of each web service. The authors have
proposed an efficient global web service selection algorithm in
order to approach quality constraints and preferences for
composite services which require aggregation of different types of
services to satisfy the user.

@@

\subsection{Web Service Community Formation}\label{sec:communities_for_rel}

>>>>

\subsection{Web Services Communities}\label{sec:CWSDefinition}
Community refers to ``the condition of sharing or having certain
attitudes and interests in common'' or ``a group of people living
in the same place or having a particular characteristic in
common''\footnote{Oxford Dictionaries}. In
\cite{DBLP:journals/internet/BenatallahSD03,
Zeng:2003:QDW:775152.775211}, the authors introduce community of
web services as collection of cooperative web services with common
functionalitiers but different QoS metrics. Therefore
$communities$ are differentiated from $composition$ types of web
service cooperation in which web services with different
functionalities work together to generate a new service with
composite functionality.

@@

Here we introduce the related research work regarding the engineering and formation
of communities of web services. In \cite{DBLP:journals/internet/BenatallahSD03}, Benatallah et al.
defined communities as \emph{Service Containers} that aggregate
substitutable web services providing a common functionality (same
set of operations). They abstracted \emph{Service Containers} as
web services that are created, advertised, discovered and invoked
just as elementary web services. The \emph{Container} is
considered as a manager that is responsible for web service
selection upon receiving a request on run-time. The authors have
proposed a scoring service based on non-functional requirements of
the request and web service capabilities to dynamically chose the
web service to perform the requested task. A similar concept was
proposed by Maamar et al. in
\cite{DBLP:journals/ijebr/MaamarSTBB09}. The authors introduced
web services communities as a collection of web services with a
common functionality but different QoS properties. A community
manager, upon receiving a request, delegates the request to one of
its current members. The choice is based on the performance
history and quality metrics of each web service. The authors have
proposed an efficient global web service selection algorithm in
order to approach quality constraints and preferences for
composite services which require aggregation of different types of
services to satisfy the user.

@@

\subsection{Web Services Community Formation}\label{sec:communities_for_rel}

### 2015-07-15 thesis/thesis/chap3/chap3.tex
changes

\subsection{Web Service Communities}\label{webservice-communities}

@@

\begin{figure*}[!t]
\centerline{\includegraphics[width=15cm]{Figures/archss.eps}}
\caption{Architecture of Web Service communities}
\label{fig_community}
\end{figure*}

@@

In this section, we present  web services and community coordinator's interactions, the task distribution process and revenue models in web service communities.

@@

The revenue function of the web service community is a linear function of $Out(C)$ with a positive constant multiplier.

@@

In this section, we present different web service community models
and focus on the problem of how both web services and community
masters as rational entities would adopt strategies to maximize
their payoff.

@@

As mentioned in Section \ref{s:tax}, a solution to help the
community stabilize is to subside the community by a relative
coefficient ($\lambda$) so the value of $\lambda v(C)$ is divided
among the community members. We have analyzed the effect of
subsiding and the cost it incurs to our web service communities.
Figure \ref{f_taxtation} shows the results. In this experiment, we
have set a community with input task rate $R_C$ of 100 and having
web services throughput rate $Th_{ws}$ values from a normal
distribution with average 10 tasks per iteration and standard
deviation 2. Part (a) shows the community size increases in a
linear fashion as ($\lambda$) increases. However, the cost (Part
(b)) is having a slight exponential growth rate since, not only
($\lambda$) increases, but also the size of the community is
increasing slowly. Therefore, subsiding can be costly for larger
number of $\lambda$ values. Part (c) depicts the number of tasks
done by the community per iteration. It is obvious that with
$\lambda$ value of 1.3, which is 30\% of the community valuation,
the number of tasks done almost reaches the input task rate cap of
100 tasks per iteration. The average quality of tasks also has a
slight increase since the community will be able to afford better
and more web services to join the community (Part (d)). These
results show the effectiveness of our subsiding method and its
impact on the QoS. In fact, using more than 30\% of the community
valuation as subsidy is not very effective and is costly to
perform.

@@

%As future work in this area, we would like to perform more analytical and theoretical analysis on the convexity condition and also minimal $\epsilon$ values in \emph{$\epsilon$-core} solution concepts based on the characteristic function in web service applications. From web service perspective, the work can be extended to consider web service compositions where a group of web services having different set of skills cooperate to perform composite tasks. Also bargaining theory from cooperating game theory concepts can be used to help web services resolve the instability and unfairness issues by side payments.

>>>>

\subsection{Web Services Communities}\label{webservice-communities}

@@

\begin{figure*}[!t]
\centerline{\includegraphics[width=15cm]{Figures/archss.eps}}
\caption{Architecture of Web Services communities}
\label{fig_community}
\end{figure*}

@@

In this section, we present  web services and community coordinator's interactions, the task distribution process and revenue models in web services communities.

@@

The revenue function of the web services community is a linear function of $Out(C)$ with a positive constant multiplier.

@@

In this section, we present different web services community models
and focus on the problem of how both web services and community
masters as rational entities would adopt strategies to maximize
their payoff.

@@

As mentioned in Section \ref{s:tax}, a solution to help the
community stabilize is to subside the community by a relative
coefficient ($\lambda$) so the value of $\lambda v(C)$ is divided
among the community members. We have analyzed the effect of
subsiding and the cost it incurs to our web services communities.
Figure \ref{f_taxtation} shows the results. In this experiment, we
have set a community with input task rate $R_C$ of 100 and having
web services throughput rate $Th_{ws}$ values from a normal
distribution with average 10 tasks per iteration and standard
deviation 2. Part (a) shows the community size increases in a
linear fashion as ($\lambda$) increases. However, the cost (Part
(b)) is having a slight exponential growth rate since, not only
($\lambda$) increases, but also the size of the community is
increasing slowly. Therefore, subsiding can be costly for larger
number of $\lambda$ values. Part (c) depicts the number of tasks
done by the community per iteration. It is obvious that with
$\lambda$ value of 1.3, which is 30\% of the community valuation,
the number of tasks done almost reaches the input task rate cap of
100 tasks per iteration. The average quality of tasks also has a
slight increase since the community will be able to afford better
and more web services to join the community (Part (d)). These
results show the effectiveness of our subsiding method and its
impact on the QoS. In fact, using more than 30\% of the community
valuation as subsidy is not very effective and is costly to
perform.

@@

%As future work in this area, we would like to perform more analytical and theoretical analysis on the convexity condition and also minimal $\epsilon$ values in \emph{$\epsilon$-core} solution concepts based on the characteristic function in web service applications. From web service perspective, the work can be extended to consider web services compositions where a group of web services having different set of skills cooperate to perform composite tasks. Also bargaining theory from cooperating game theory concepts can be used to help web services resolve the instability and unfairness issues by side payments.

### 2015-07-15 thesis/thesis/chap4/chap4.tex
changes

Web Service communities have the advantages of facilitating web service discovery and providing better quality of service compared to individual services. Communities act as abstract web services, communicating with external entities via the same standard protocols that a normal web service employs. The difference is that communities regulate the service process via sophisticated internal communication protocols, thereby providing services based on the combined efforts of a number of web services. The downside to communities is the complexity of management involved in finding and inviting adequate individual services and managing the overall quality of the combined work of several services.
% because although they have similar functionality, they have different attitudes.
When interacting with a community of web services, users send their requests to the coordinator of the community, which plays the role of community representative or access point. The community coordinator is responsible of receiving tasks and delivering services. Moreover, as community representative, it verifies the credentials of new web services before accepting them into the community and kicks services that could harm the value of the community.

@@

\begin{figure}%[!t]
\centering
\includegraphics[width=3.5in]{figures/utility_gain.eps}
\caption{DDM against Rational: utility gain }
\label{utility_gain_mlisa_and_rational}
\end{figure}

\begin{figure}%[!t]
\centering
\includegraphics[width=3.5in]{figures/utility_ratio.eps}
\caption{DDM against Rational: ratio of utility gain}
\label{utility_gain_mlisa_and_rational_ratio}
\end{figure}

@@

\begin{figure}%[!t]
\centering
\includegraphics[width=3.5in]{figures/roc.eps}
\caption{RoC Curve}
\label{roc5}
\end{figure}

@@

\begin{figure}%[!t]
\centering
\includegraphics[width=3.5in]{figures/stats1.eps}
\caption{Overall utility of all the communities}
\label{stats1}
\end{figure}

\begin{figure}%[!t]
\centering
\includegraphics[width=3.5in]{figures/stats2.eps}
\caption{Utility gain over time}
\label{stats2}
\end{figure}

@@

\begin{figure}%[!t]
\centering
\includegraphics[width=3.5in]{figures/stats3.eps}
\caption{Average community size}
\label{stats3}
\end{figure}

@@

%\noindent \textbf{Web Service Communities}

@@

%Here we introduce the related research work regarding the engineering and formation of communities of web services. In \cite{DBLP:journals/internet/BenatallahSD03}, Benatallah et al. defined communities of web services as \emph{Service Containers} that aggregate substitutable web services having the same set of operations and providing common functionalities. They abstracted \emph{Service Containers} as web services that are created, advertised, discovered and invoked just as elementary web services. The \emph{Container} is considered as a manager that is responsible for web service selection upon receiving a request on run-time. The authors have proposed a scoring technique based on non-functional requirements of the request and web service capabilities to dynamically chose the web service to perform the requested task. A similar concept was proposed by Maamar et al. in \cite{DBLP:journals/ijebr/MaamarSTBB09}. The authors introduced web service communities as a collection of web services with a common functionality but different QoS properties. A community manager, upon receiving a request, delegates the request to one of its current members. The choice is based on the performance history and quality metrics of each web service. The authors have proposed an efficient global web service selection algorithm in order to approach quality constraints and preferences for composite services which requires aggregation of different types of services to satisfy the user. In the same line of research, Benslimane et al. \cite{Liris-2770} have proposed a multi-layer approach grouping similar web services into communities and having an interface implemented as an abstract web service for accessing the community. The considered layers are: composite, management and community. The interactions among those layers and the bindings are performed by a generic driver called Open Software Connectivity (OSC). However, unlike our work, these proposals only tackle the problem of internal community management from the task execution perspective, but not the problem of community formation and the one of dynamically selecting which community to join and which web service to invite. Moreover, the web services selection process is done centrally by the community manager, while we address the problem more realistically from the distributed angle.

>>>>

Web Services communities have the advantages of facilitating web service discovery and providing better quality of service compared to individual services. Communities act as abstract web services, communicating with external entities via the same standard protocols that a normal web service employs. The difference is that communities regulate the service process via sophisticated internal communication protocols, thereby providing services based on the combined efforts of a number of web services. The downside to communities is the complexity of management involved in finding and inviting adequate individual services and managing the overall quality of the combined work of several services.
% because although they have similar functionality, they have different attitudes.
When interacting with a community of web services, users send their requests to the coordinator of the community, which plays the role of community representative or access point. The community coordinator is responsible of receiving tasks and delivering services. Moreover, as community representative, it verifies the credentials of new web services before accepting them into the community and kicks services that could harm the value of the community.

@@

\begin{figure}%[!t]
\centering
\includegraphics[width=5in]{figures/utility_gain.eps}
\caption{DDM against Rational: utility gain }
\label{utility_gain_mlisa_and_rational}
\end{figure}

\begin{figure}%[!t]
\centering
\includegraphics[width=5in]{figures/utility_ratio.eps}
\caption{DDM against Rational: ratio of utility gain}
\label{utility_gain_mlisa_and_rational_ratio}
\end{figure}

@@

\begin{figure}%[!t]
\centering
\includegraphics[width=4in]{figures/roc.eps}
\caption{RoC Curve}
\label{roc5}
\end{figure}

@@

\begin{figure}%[!t]
\centering
\includegraphics[width=4in]{figures/stats1.eps}
\caption{Overall utility of all the communities}
\label{stats1}
\end{figure}

\begin{figure}%[!t]
\centering
\includegraphics[width=4in]{figures/stats2.eps}
\caption{Utility gain over time}
\label{stats2}
\end{figure}

@@

\begin{figure}%[!t]
\centering
\includegraphics[width=4in]{figures/stats3.eps}
\caption{Average community size}
\label{stats3}
\end{figure}

@@

%\noindent \textbf{Web Services Communities}

@@

%Here we introduce the related research work regarding the engineering and formation of communities of web services. In \cite{DBLP:journals/internet/BenatallahSD03}, Benatallah et al. defined communities of web services as \emph{Service Containers} that aggregate substitutable web services having the same set of operations and providing common functionalities. They abstracted \emph{Service Containers} as web services that are created, advertised, discovered and invoked just as elementary web services. The \emph{Container} is considered as a manager that is responsible for web service selection upon receiving a request on run-time. The authors have proposed a scoring technique based on non-functional requirements of the request and web service capabilities to dynamically chose the web service to perform the requested task. A similar concept was proposed by Maamar et al. in \cite{DBLP:journals/ijebr/MaamarSTBB09}. The authors introduced web services communities as a collection of web services with a common functionality but different QoS properties. A community manager, upon receiving a request, delegates the request to one of its current members. The choice is based on the performance history and quality metrics of each web service. The authors have proposed an efficient global web service selection algorithm in order to approach quality constraints and preferences for composite services which requires aggregation of different types of services to satisfy the user. In the same line of research, Benslimane et al. \cite{Liris-2770} have proposed a multi-layer approach grouping similar web services into communities and having an interface implemented as an abstract web service for accessing the community. The considered layers are: composite, management and community. The interactions among those layers and the bindings are performed by a generic driver called Open Software Connectivity (OSC). However, unlike our work, these proposals only tackle the problem of internal community management from the task execution perspective, but not the problem of community formation and the one of dynamically selecting which community to join and which web service to invite. Moreover, the web services selection process is done centrally by the community manager, while we address the problem more realistically from the distributed angle.

### 2015-07-15 thesis/thesis/chap6/chap6.tex
changes

As future work, for the community formation in our first model, we would like to perform more analytical and theoretical analysis on the convexity condition and also minimal
$\epsilon$ values in \emph{$\epsilon$-core} solution concepts based on the characteristic function in web service applications. From web service perspective, the work can be extended to consider web service compositions where a group of web services having different set of skills cooperate to perform composite tasks. Also bargaining theory from cooperating game theory concepts can be used to help web services resolve the instability and unfairness issues by side payments.

>>>>

As future work, for the community formation in our first model, we would like to perform more analytical and theoretical analysis on the convexity condition and also minimal
$\epsilon$ values in \emph{$\epsilon$-core} solution concepts based on the characteristic function in web service applications. From web service perspective, the work can be extended to consider web services compositions where a group of web services having different set of skills cooperate to perform composite tasks. Also bargaining theory from cooperating game theory concepts can be used to help web services resolve the instability and unfairness issues by side payments.
